\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-30 OaklandSubmission/conclude.tex
cutting

\section{Conclusions}
\label{sec:conclude}

Therefore, we believe
that \tool generated code can be a better baseline than Yao-based 2PC
when comparing against generic 2PC-based implementations for secure
prediction tasks. \aseem{I don't understand this last sentence. Why is
  ours a better baseline?} \nc{This last line is more a comment for a ``conclusion'' and is a bit out of place here. I guess we want to say that till recently if generic 2PC using Yao was not feasible, then people had to come up with specialized protocols and didnt have any benchmark or had a poor benchmark (of Yao) to compare their 2PC with. Moving forward, our language generated 2PC can be the baseline benchmark for all 2PC protocols.}

>>>>

\section{Conclusion}
\label{sec:conclude}
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs.
The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous works

### 2017-10-30 OaklandSubmission/eval.tex
cutting

We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$. We focus on the machine learning models for {\it classification}. A classifier $C$ uses a trained model to {\it predict} a
label $\ell$ for an input data point $x$. For example, given a
data point which is a tuple of humidity and temperature 
a classifier might be able to predict a label as ``will rain'' or ``will not rain''. The {\it
  accuracy} of a classifier refers to the fraction of data points that
the classifier labels correctly from a given set of test data
points.

\subsubsection*{Standard classifiers}
A {\em binary linear classifier} is one of the simplest classifiers. Here,
the input is a data point $x\in\R^d$,
and the model is a vector $w\in\R^d$.
 The possible labels are
$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
$C_w\equiv w^T\matmul x>0$, where $w^T$ denotes the transpose
of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
This classifier requires $d$ multiplications, $d-1$
additions, and a single comparison.

A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
from the set $[n]$~\cite{shafindss}.
Here, the input data point is a {\it feature}
vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
%The model has two matrices: a vector $P$ of length $n$ s.t. for each
%$i
%\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
%that the output label is $i$ for $i\in[n]$.
%The other matrix $T$ has size $d\times F\times n$ and for each $j\in
%[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
%\equiv\log p(x_j|\ell=i)$, i.e.,
%the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
%The classifier $C_{P,T}(x)$ outputs
%\[
%{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
%\]
%The additions in the expression above sum $n$-dimensional vectors.
%The model size of this classifier is $|P|+|T|$.
This classifier requires $\Theta(ndF)$ comparisons and additions. %Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
  %Section~\ref{sec:impl}). %\nc{Made DG's comment as part of regular text - worth point out}. \cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
%
A {\em decision tree} of size $N$ takes as input an $x\in\R^d$
%,
 and 
%the
%model consists of a binary tree of depth $d$ with a boolean predicate
%assigned to each internal node.
% The root note is the $0^{th}$ node and
%for an internal node $j \in [N]$, the children nodes are $2j+1$ and
%$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
%$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
%root and if the predicate at the current node $b_j^i$ is false (resp.,
%true) then $x$ is passed to the left (resp., right) child with
%predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
%repeated till we reach a leaf. The leaves of the tree are labels and
%the output label is the leaf visited by this traversal.
%Such a (binary) decision tree 
%and can be encoded as a polynomial (of
%degree linear in depth of the tree) and then
 the prediction task
reduces to polynomial evaluation~\cite{shafindss}.
% E.g., a binary tree with depth one
%and size $N=3$ can be encoded as the polynomial
%$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
%root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
%of the right leaf.

@@

Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
and is zero otherwise. 
The operation $\circ$ is a pointwise multiplication of two vectors, $W_j$'s and $V_j$'s
are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix). \bonsai can be
seen as a variant of decision trees where the prediction is a function
of the path traversed from root to the leaf and not just the leaf
itself.

@@

\subsection{Standard classifiers}
\label{sec:shallow}
We train (on cleartext data) the three standard classifiers, linear, Na\"{i}ve Bayes, and
decision trees, on the following data sets from the UCI
machine learning repository~\cite{uci}:
 the Wisconsin Breast Cancer data set, 
Credit Approval data set, Audiology (Standardized) data set, Nursery
data set, and ECG (electrocardiogram) classification data
from~\cite{barni}.

\subsubsection*{Linear classifier}The results for linear
classification are in Table~\ref{tab:lc}. The
input and the model are both vectors of length $d$. The columns
``Prev. time'' and ``Prev. comm'' show the time and the total
network communication reported by Bost et al.~\cite{shafindss} for a
network setting with 40ms
round trip time, which is same as our WAN setting. The total execution
time
of \tool generated code in the LAN and the WAN setting is reported
next, followed by the total communication.
We observe that the \tool code performance matches the hand-crafted protocol of Bost
et al., and the programmer effort in \tool is just 20
lines (last column in the table) of high-level code in the \tool
source language. %\aseem{They write circuits,
 % right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).} 
% Our execution time is dominated by the multiplications and
%boolean-ands.
%The number of boolean-and gates (\#And) remains constant for both the
%benchmarks as they reflect the number of gates required for a
%single  comparison, while the number of
%multiplication gates (\#Mul) is the same as the model size $d$.
%% The
%% last column shows the lines of code of \tool program, which is also
%% independent of the model size.

@@

\subsubsection*{Na\"{i}ve Bayes} The results for Na\"{i}ve Bayes are
in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
feature vector $x$ and $F$ is the number of possible values of $x_i$
for each $i \in [d]$.
Although, there are no multiplications in these benchmarks, these do
have a significant number of comparisons (due to array look-ups using
secret indices) that
result in many boolean-and gates. %(13k/750k denote 13000/75000). 
As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
show the numbers reported
by Bost et al.~\cite{shafindss}. We observe that \tool generated code
has better performance, despite using a generic 2PC,
as opposed to custom designed protocols developed by Bost et
al. Moreover, they remark that in their setup, generic Yao-based 2PC
did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
they had to scale down the prediction task, and even then their
implementation was 500x slower. Whereas, we show that by using a
cryptographic-cost aware compiler, we can scale generic 2PC to real
prediction tasks, and get performance competitive to or better than the
specialized protocols.

@@

\subsubsection*{Decision trees}
Our decision tree numbers in Table~\ref{tab:dt} further validate
this claim. We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
describe specialized protocols for trees and forests (that perform
better than~\cite{shafindss}).
Similar to Bost et al.~\cite{shafindss}, they also remarks that these
computations are infeasible with Yao-based 2PC.
Again, we observe that
the performance of \tool generated generic 2PC
code is competitive with specialized protocols.

@@

%dataset | time | comm | timeL | timeW | comm
\subsection{Deep neural nets}
\label{sec:dnneval}
We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
comparison, we consider their implementations from
\minion~\cite{minionn}, which outperforms their previous
implementations. Table~\ref{tab:nn} shows the
results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
We note that for each of these DNNs, \minion provides a
specialized protocol, while \tool uses a generic 2PC protocol
(auto) generated from high-level code.

@@

In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
The column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
We note that \minion and \tool implementations are both based on ABY.
These results are surprising as \minion has a much more efficient
preprocessing phase and has SIMD 
(single instruction multiple data) capabilities that allow it to perform
matrix operations efficiently. In contrast, \tool focuses on generic
2PC, and such optimizations are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in Section~\ref{sec:pipeeval}.
%net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length

@@

\subsection{State-of-the-art classifiers}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
vector.
We implement this classifer in \tool and present the results in the
first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.

@@

 \caption{Bonsai benchmarks}
 \label{tab:bonsai} 
\end{table}
\setlength\tabcolsep{6pt}
\subsection{Partitioning}
\label{sec:pipeeval}
We evaluate partitioning next.
\subsubsection*{CIFAR-10}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchamrk is 336 lines.

Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth (the heaviest) stage. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
For this DNN, the specialized \minion protocol takes 544 seconds and
communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to arbitrary sized
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed. 
\begin{table}
\begin{tabular}{|c|c|c|c |c|c|c|}
\hline
           &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates \\
\hline
Total      &  265.6       & 647.5        & 40683       & 21m    & 61m    &  337m  \\
\hline
Stage 6    &  55.2        & 122.6        & 6744        & 12m    & 10m   &  98m  \\
\hline
\end{tabular}

@@

>>>>

We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$. We focus on the machine learning models for {\it classification}. A classifier $C$ uses a trained model to {\it predict} a
label $\ell$ for an input data point $x$. For example, given a
data point which is a tuple of humidity and temperature 
a classifier might be able to predict a label ``will rain'' or ``will not rain''.
The {\it model size} of a classifier is the number of parameters in the model.
For example, the model size of the classifier in~\figureref{ex-sml} is $|w|+1=31$.
The {\it
  accuracy} of a classifier refers to the fraction of data points that
the classifier labels correctly from a given set of test data
points.

\subsubsection*{Standard classifiers}
A {\em binary linear classifier} is one of the simplest classifiers. Here,
the input is a data point $x\in\R^d$,
and the model is a vector $w\in\R^d$.
 The possible labels are
$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
$C_w\equiv w^Tx>0$.
%, where $w^T$ denotes the transpose
%of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns %$\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
%The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
%This classifier requires $d$ multiplications, $d-1$
%additions, and a single comparison.
%
A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
from the set $[n]$~\cite{shafindss}.
Here, the input data point is a {\it feature}
vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
%The model has two matrices: a vector $P$ of length $n$ s.t. for each
%$i
%\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
%that the output label is $i$ for $i\in[n]$.
%The other matrix $T$ has size $d\times F\times n$ and for each $j\in
%[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
%\equiv\log p(x_j|\ell=i)$, i.e.,
%the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
%The classifier $C_{P,T}(x)$ outputs
%\[
%{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
%\]
%The additions in the expression above sum $n$-dimensional vectors.
%The model size of this classifier is $|P|+|T|$.
This classifier requires $\Theta(ndF)$ comparisons and additions. %Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
  %Section~\ref{sec:impl}). %\nc{Made DG's comment as part of regular text - worth point out}. \cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
%
A {\em decision tree} of size $N$ takes as input an $x\in\R^d$
%,
 and 
%the
%model consists of a binary tree of depth $d$ with a boolean predicate
%assigned to each internal node.
% The root note is the $0^{th}$ node and
%for an internal node $j \in [N]$, the children nodes are $2j+1$ and
%$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
%$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
%root and if the predicate at the current node $b_j^i$ is false (resp.,
%true) then $x$ is passed to the left (resp., right) child with
%predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
%repeated till we reach a leaf. The leaves of the tree are labels and
%the output label is the leaf visited by this traversal.
%Such a (binary) decision tree 
%and can be encoded as a polynomial (of
%degree linear in depth of the tree) and then
 the prediction task
reduces to polynomial evaluation~\cite{shafindss}.
% E.g., a binary tree with depth one
%and size $N=3$ can be encoded as the polynomial
%$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
%root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
%of the right leaf.

@@

Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
and is zero otherwise. 
The operation $\circ$ is a pointwise multiplication of two vectors, $W_j$'s and $V_j$'s
are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix).
%\bonsai can be
%seen as a variant of decision trees where the prediction is a function
%of the path traversed from root to the leaf and not just the leaf
%itself.

@@

\subsection{Secure prediction}
\noindent\textbf{Standard classifiers}
%\label{sec:shallow}
We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
decision trees, from~\cite{shafindss} on the following data sets from the UCI
machine learning repository~\cite{uci}:
 the Wisconsin Breast Cancer data set, 
Credit Approval data set, Audiology (Standardized) data set, Nursery
data set, and ECG (electrocardiogram) classification data
from~\cite{barni}.

The results for linear
classification are in Table~\ref{tab:lc}.
The input and the model are both vectors of length $d$. 
The columns ``Prev. time'' and ``Prev. comm'' show the time and the total
network communication reported by Bost et al.~\cite{shafindss} for a
network setting with 40ms
round trip time, which is same as our WAN setting. The total execution
time
of \tool generated code in the LAN and the WAN setting is reported
next, followed by the total communication.
We observe that the \tool code performance matches the hand-crafted protocol of Bost
et al., and the programmer effort in \tool is just 20
lines (last column in the table) of high-level code in the \tool
source language. %\aseem{They write circuits,
 % right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).} 
% Our execution time is dominated by the multiplications and
%boolean-ands.
%The number of boolean-and gates (\#And) remains constant for both the
%benchmarks as they reflect the number of gates required for a
%single  comparison, while the number of
%multiplication gates (\#Mul) is the same as the model size $d$.
%% The
%% last column shows the lines of code of \tool program, which is also
%% independent of the model size.

@@

%\subsubsection*{Na\"{i}ve Bayes}

@@

%\subsubsection*{Decision trees}

@@

%dataset | time | comm | timeL | timeW | comm
%Our decision tree numbers in Table~\ref{tab:dt} further validate
%this claim. 
%We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
%describe specialized protocols for trees and forests (that perform
%better than~\cite{shafindss}).
%Similar to Bost et al.~\cite{shafindss}, they also remarks that these
%computations are infeasible with Yao-based 2PC.
%Again, we observe that
%the performance of \tool generated generic 2PC
%code is competitive with specialized protocols. 
The results for Na\"{i}ve Bayes are
in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
feature vector $x$ and $F$ is the number of possible values of $x_i$
for each $i \in [d]$.
%Although, there are no multiplications in these benchmarks, these do
%have a significant number of comparisons (due to array look-ups using
%secret indices) that
%result in many boolean-and gates. %(13k/750k denote 13000/75000). 
As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
show the numbers reported
by Bost et al.~\cite{shafindss}. We observe that \tool generated code
has better performance, despite using a generic 2PC,
as opposed to custom designed protocols developed by Bost et
al. Moreover, they remark that in their setup, generic Yao-based 2PC
did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
they had to scale down the prediction task, and even then their
implementation was 500x slower. Whereas, we show that by using a
cryptographic-cost aware compiler, we can scale generic 2PC to real
prediction tasks, and get performance competitive to or better than the
specialized protocols. Table~\ref{tab:dt} 
compares against the more recent work of~\cite{wu}
on decision trees and further validates this claim.
\\\\
\noindent\textbf{Deep neural nets}
We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
comparison, we consider their implementations from
\minion~\cite{minionn}, which outperforms their previous
implementations. Table~\ref{tab:nn} shows the
results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
We note that for each of these DNNs, \minion provides a
specialized protocol, while \tool uses a generic 2PC protocol
(auto) generated from high-level code.

@@

%net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length

@@

In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
The column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
We note that \minion and \tool implementations are both based on ABY.
These results are surprising as \minion has a much more efficient
preprocessing phase and has SIMD 
(single instruction multiple data) capabilities that allow it to perform
matrix operations efficiently. In contrast, \tool focuses on generic
2PC, and such optimizations are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in Section~\ref{sec:pipeeval}.
\\\\
\noindent\textbf{State-of-the-art classifiers}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of
digits 0 to 9.
The first model that the tutorial describes is a softmax regression
that provides an accuracy of 92\%. The classifier evaluates
$ \mathsf{argmax}\ W\matmul x+b$.
Here, $x$ is a 784 length vector obtained from the input image,
each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
vector.
We implement this classifer in \tool and present the results in the
first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.

@@

 \caption{Bonsai benchmarks}
 \label{tab:bonsai} 
\end{table}
\setlength\tabcolsep{6pt}
\subsection{Secure code partitioning}
\label{sec:pipeeval}
\noindent\textbf{CIFAR-10}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchamrk is 336 lines.

\begin{table}
\begin{tabular}{|c|c|c|c |c|c|c|}
\hline
           &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates \\
\hline
Total      &  265.6       & 647.5        & 40683       & 21m    & 61m    &  337m  \\
\hline
Stage 6    &  55.2        & 122.6        & 6744        & 12m    & 10m   &  98m  \\
\hline
\end{tabular}

@@

Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth (the heaviest) stage. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
For this DNN, the specialized \minion protocol takes 544 seconds and
communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to arbitrary sized
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed.
\\\\
\noindent\textbf{Matrix factorization}
\tool, as a language, is not tied to secure prediction and can express more general computations.
To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.

\begin{table}
\begin{tabular}{|c|c|c|c |c|c| c|}
\hline
  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
\hline
1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
\hline
2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
\hline
3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
\hline
Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
\hline
\end{tabular}

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
 \label{tab:factor} 
\end{table}

We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
However, this benchmark requires more lines of code than the previous benchmarks
because of  Batcher's sort (450 lines of \tool code in each stage).
However, the current programmer effort seems miniscule compared to the mammoth implementation effort
put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
crypto-backend to this benchmark.

### 2017-10-30 OaklandSubmission/impl.tex
cutting

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to bound check the array indices
($P\vDash e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}})),
Figure~\ref{fig:compile}). We take the \tool source program and
translate it as an input to the solver. The solver takes less
than a minute on our largest benchmark to verify that all the array
accesses are in-bounds.

@@

The compilation rules of Figure~\ref{fig:compile} can introduce
repeated coercions from arithmetic to
boolean and vice versa.
The \tool implementation performs  the standard ``common subexpression elimination"
optimization~\cite{dragonbook} to reduce these coercions.
On each coercion, \tool memorizes the pair of arithmetic
share and boolean share involved in the coercion. 
\tool invalidates such pairs when the variables corresponding to the
shares are overwritten by assignments. 
In subsequent coercions, \tool  reuses valid pairs (if available)
instead of inserting code to recompute them afresh.

>>>>

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to bound check the array indices
($\models e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}})),
Figure~\ref{fig:compile}). We take the \tool source program and
translate it as an input to the solver. The solver takes less
than a minute on our largest benchmark to verify that all the array
accesses are in-bounds.

@@

The compilation rules of Figure~\ref{fig:compile} can introduce
repeated coercions from arithmetic to
boolean and vice versa.
Since \tool is aware of the cryptographic costs associated with these coercions,
it tries to minimize them using several optimizations.
For instance, if boolean shares of a variable are available in scope and the variable is involved in
an addition then \tool performs the addition in boolean rather than first coercing boolean to arithmetic
and then performing an arithmetic addition. However, multiplication is always performed in arithmetic
as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
Other optimizations include the standard ``common subexpression elimination"
optimization~\cite{dragonbook}.
On each coercion, \tool memorizes the pair of arithmetic
share and boolean share involved in the coercion. 
\tool invalidates such pairs when the variables corresponding to the
shares are overwritten by assignments. 
In subsequent coercions, \tool  reuses valid pairs (if available)
instead of inserting code to recompute them afresh.
